Manuscript 1 code

### 2024-09-13 manuscript_1_code.Rmd

```{r}
(figure_2b <- ggplot() +
  geom_point(data = site_richness_df, aes(x = year, y = richness_mean, colour = habitat), size = 3) +
  geom_line(data = site_richness_df, aes(x = year, y = richness_mean, colour = habitat), linewidth = 1) +
  geom_errorbar(data = site_richness_df, aes(x = year, ymin = richness_mean - richness_se, ymax = richness_mean + richness_se, colour = habitat), size = 1.3, width = 0) +
  scale_colour_manual(values = habitat_colours) +
  #geom_segment(aes(y = 75, yend = 75, x = 2006, xend = 2010), linewidth = 2, colour = "grey") +
  annotate("segment", x = 2006, xend = 2010, y = 11, yend = 11, linewidth = 2, colour = "grey") +
  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
   scale_x_continuous(breaks = unique(site_richness_df$year)) +
   labs(y = "Macroalgal Richness \n",
        x = "Year",
        title = "b. Site-level Richness") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title.y = element_text(size=18), 
        plot.title = element_text(size=18),
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_rect(fill = "white"),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_blank(),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black")) )

@@

## 2c Evenness @ Quad by Year

```{r calculate evennes}
#dear callie, why is this here?
site_richness_df <- alpha_diversity_site_macro %>% 
  group_by(year, habitat) %>% 
  summarise(richness_mean = mean(richness),
            richness_se = se(richness))

@@

```{r}
(figure_2c <- ggplot() +
  geom_point(data = quad_evenness_df, aes(x = year, y = evenness_mean, colour = habitat), size = 3) +
  geom_line(data = quad_evenness_df, aes(x = year, y = evenness_mean, colour = habitat), linewidth = 1) +
  geom_errorbar(data = quad_evenness_df, aes(x = year, ymin = evenness_mean - evenness_se, ymax = evenness_mean + evenness_se,
                                             colour = habitat), width = 0, size = 1.3) +
  scale_colour_manual(values = habitat_colours) +
  # scale_fill_manual(values = habitat_colours) +
  #geom_segment(aes(y = 75, yend = 75, x = 2006, xend = 2010), linewidth = 2, colour = "grey") +
  annotate("segment", x = 2006, xend = 2010, y = .7, yend = .7, linewidth = 2, colour = "grey") +
  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
   scale_x_continuous(breaks = unique(quad_evenness_df$year)) +
   labs(y = "Macroalgal evenness \n",
        x = "Year",
        title = "c. Quadrat-level evenness") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title.y = element_text(size=18), 
        plot.title = element_text(size=18),
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_rect(fill = "white"),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_blank(),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black")) )

@@

## 2d Evenness @ Site by Year

```{r set up evenness}
site_evenness_df <- alpha_diversity_site_macro %>% 
  drop_na(evenness) %>%
  group_by(year, habitat) %>% 
  summarise(evenness_mean = mean(evenness),
            evenness_se = se(evenness))

@@

```{r plot evenness site}
(figure_2d <- ggplot() +
  geom_point(data = site_evenness_df, aes(x = year, y = evenness_mean, colour = habitat), size = 3) +
  geom_line(data = site_evenness_df, aes(x = year, y = evenness_mean, colour = habitat), linewidth = 1) +
  geom_errorbar(data = site_evenness_df, aes(x = year, ymin = evenness_mean - evenness_se, ymax = evenness_mean + evenness_se, colour = habitat), size = 1.3, width = 0) +
  scale_colour_manual(values = habitat_colours) +
  #geom_segment(aes(y = 75, yend = 75, x = 2006, xend = 2010), linewidth = 2, colour = "grey") +
  annotate("segment", x = 2006, xend = 2010, y = 1, yend = 1, linewidth = 2, colour = "grey") +
  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
   scale_x_continuous(breaks = unique(site_evenness_df$year)) +
   labs(y = "Macroalgal evenness \n",
        x = "Year",
        title = "d. Site-level evenness") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title.y = element_text(size=18), 
        plot.title = element_text(size=18),
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_rect(fill = "white"),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_blank(),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black")) )

>>>>

### plot
```{r}
(figure_2b <- ggplot() +
  geom_point(data = site_richness_df, aes(x = year, y = richness_mean, colour = habitat), size = 3) +
  geom_line(data = site_richness_df, aes(x = year, y = richness_mean, colour = habitat), linewidth = 1) +
  geom_errorbar(data = site_richness_df, aes(x = year, ymin = richness_mean - richness_se, ymax = richness_mean + richness_se, colour = habitat), size = 1.3, width = 0) +
  scale_colour_manual(values = habitat_colours) +
  #geom_segment(aes(y = 75, yend = 75, x = 2006, xend = 2010), linewidth = 2, colour = "grey") +
  annotate("segment", x = 2006, xend = 2010, y = 11, yend = 11, linewidth = 2, colour = "grey") +
  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
   scale_x_continuous(breaks = unique(site_richness_df$year)) +
   labs(y = "Macroalgal Richness \n",
        x = "Year",
        title = "b. Site-level Richness") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title.y = element_text(size=18), 
        plot.title = element_text(size=18),
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_rect(fill = "white"),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_blank(),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black")) )

@@

## 2c Evenness @ Quad by Year
### prep
```{r calculate evennes}
#dear callie, why is this here?
site_richness_df <- alpha_diversity_site_macro %>% 
  group_by(year, habitat) %>% 
  summarise(richness_mean = mean(richness),
            richness_se = se(richness))

@@

### plot
```{r}
(figure_2c <- ggplot() +
  geom_point(data = quad_evenness_df, aes(x = year, y = evenness_mean, colour = habitat), size = 3) +
  geom_line(data = quad_evenness_df, aes(x = year, y = evenness_mean, colour = habitat), linewidth = 1) +
  geom_errorbar(data = quad_evenness_df, aes(x = year, ymin = evenness_mean - evenness_se, ymax = evenness_mean + evenness_se,
                                             colour = habitat), width = 0, size = 1.3) +
  scale_colour_manual(values = habitat_colours) +
  # scale_fill_manual(values = habitat_colours) +
  #geom_segment(aes(y = 75, yend = 75, x = 2006, xend = 2010), linewidth = 2, colour = "grey") +
  annotate("segment", x = 2006, xend = 2010, y = .7, yend = .7, linewidth = 2, colour = "grey") +
  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
   scale_x_continuous(breaks = unique(quad_evenness_df$year)) +
   labs(y = "Macroalgal evenness \n",
        x = "Year",
        title = "c. Quadrat-level evenness") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title.y = element_text(size=18), 
        plot.title = element_text(size=18),
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_rect(fill = "white"),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_blank(),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black")) )

@@

## 2d Evenness @ Site by Year
### prep
```{r set up evenness}
site_evenness_df <- alpha_diversity_site_macro %>% 
  drop_na(evenness) %>%
  group_by(year, habitat) %>% 
  summarise(evenness_mean = mean(evenness),
            evenness_se = se(evenness))

@@

### plot
```{r plot evenness site}
(figure_2d <- ggplot() +
  geom_point(data = site_evenness_df, aes(x = year, y = evenness_mean, colour = habitat), size = 3) +
  geom_line(data = site_evenness_df, aes(x = year, y = evenness_mean, colour = habitat), linewidth = 1) +
  geom_errorbar(data = site_evenness_df, aes(x = year, ymin = evenness_mean - evenness_se, ymax = evenness_mean + evenness_se, colour = habitat), size = 1.3, width = 0) +
  scale_colour_manual(values = habitat_colours) +
  #geom_segment(aes(y = 75, yend = 75, x = 2006, xend = 2010), linewidth = 2, colour = "grey") +
  annotate("segment", x = 2006, xend = 2010, y = 1, yend = 1, linewidth = 2, colour = "grey") +
  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
   scale_x_continuous(breaks = unique(site_evenness_df$year)) +
   labs(y = "Macroalgal evenness \n",
        x = "Year",
        title = "d. Site-level evenness") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title.y = element_text(size=18), 
        plot.title = element_text(size=18),
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_rect(fill = "white"),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_blank(),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black")) )
